---
title: "Survival Analyses Exercises - Day 4"
author: Emin Onur Karakaslar
output: md_document
---


### Fit a Cox model

```{r load libs}
library(dplyr)
library(haven)
library(survival)
library(survminer)
```
#### 1.
```{r}
df <- read_sav("Day 4_pbc.sav")

fit.dpca.cox <- coxph(Surv(years, status) ~ rx + bilirubin, data = df)

fit.dpca.cox
```



#### 2. 

```{r}
fit.dpca.cox.histology <- coxph(Surv(years, status) ~ as.factor(histol), data = df)
fit.dpca.cox.histology
```




#### 3.
```{r}
fit.dpca.cox.age <- coxph(Surv(years, status) ~ age, data = df)
fit.dpca.cox.age
```

```{r}
# lower CI
exp(0.039995 - 0.008811 * 1.96)
```
```{r}
# upper CI
exp(0.039995 + 0.008811 * 1.96)
```

#### 4. 
```{r}
fit.dpca.cox.age <- coxph(Surv(years, status) ~ age, data = df)
fit.dpca.cox.age
```

```{r}
exp(0.039995) ^ 5
```
```{r}
#lower CI
exp(0.039995 - 1.96 * 0.008811) ^ 5
```

```{r}
#upper CI
exp(0.039995+ 1.96 * 0.008811) ^ 5
```

#### 5.
```{r}
fit.dpca.cox.bilirubin <- coxph(Surv(years, status) ~ bilirubin, data = df)
fit.dpca.cox.bilirubin
```

```{r}
quantile(df$bilirubin)
```
```{r}
fit.dpca.cox.bilirubin2 <- coxph(Surv(years, status) ~ bilirubin + hepatom + spiders + edema, data = df)
fit.dpca.cox.bilirubin2
```

### Interaction

#### 1.
```{r}
fit.dpca.rxhepa <- coxph(Surv(years, status) ~ rx * hepatom, data = df)
fit.dpca.rxhepa
```

#### 2.
No, as p = 0.81

#### 3.



### Predicted survival for specific covariate patterns

#### 1.
```{r}
fit.dpca.bilhep <- coxph(Surv(years, status) ~ bilirubin + hepatom, data = df, x = T)
fit.dpca.bilhep
```

```{r}
head(df[,c("hepatom","bilirubin")])
```
```{r}
df.hepbil <- df %>% group_by(hepatom) %>% summarise(bilirubin = mean(bilirubin))
```

```{r}
library(pec)

plotPredictSurvProb(fit.dpca.bilhep, newdata = df.hepbil, 
                    col = c("green", "blue"), legend = T)
legend("bottomleft", legend=c("hepatom=0", "hepatom=1"),
       lty=c(1,1),col=c("green", "blue"))
```
#### 2.
No you can't estimate mean survival looking at these graphs as these are predictions.


### Stratified Cox Model and test for the proportional hazard assumption

#### 1.
```{r}
fit.KM <- survfit(Surv(years, status) ~ edema, data = df)

ggsurvplot(fit.KM, conf.int = T, pval = T, risk.table = T)
```
```{r}
ggsurvplot(fit.KM, conf.int = T, pval = T, risk.table = T, fun = "cumhaz")
```
```{r}
fit.KM
```


Looking at the KM, proportional hazard assumption holds as there is no crossing point between lines.

#### 2.
```{r}
fit.cox.edema <- coxph(Surv(years, status) ~ edema + age, data = df, x = T)
fit.cox.edema
```

```{r}
ggsurvplot(survfit(fit.cox.edema), data = df, palette = c("#2E9FDF"))
```

```{r}
df_new <- with(df, data.frame(edema = c(0,1), age = rep(mean(age, na.rm = TRUE), 2)))

df_new
fit.new <- survfit(fit.cox.edema, newdata = df_new)
ggsurvplot(fit.new, data = df_new)
```
This seems like a better KM plot compared to earlier version.



```{r}
fit.cox.edema.strata <- coxph(Surv(years, status) ~ strata(edema) + age, data = df, x = T)
fit.cox.edema.strata
```

