---
title: "Survival Analyses Exercises - Day 2"
author: Emin Onur Karakaslar
output: md_document
---


```{r load libs}
library(haven)
library(survival)
library(survminer)
```

```{r}
df <- read_sav("Day 2_Logrank_example.sav")

head(df)
```

```{r}
fit.KM <- survfit(Surv(time, status) ~ group, data = df)

summary(fit.KM)
```

```{r}
survdiff(Surv(time, status) ~ group, data = df)
```



```{r}
ggsurvplot(fit.KM, conf.int = T, pval = T)
```

### Minimum, maximum and median follow-up
```{r}
df2 <- read_sav("Day 2_drug6mp.sav")
```


finding the median follow-up
```{r median follow-up}
df2$followup <- 1 - df2$Status

fit.KM2 <- survfit(Surv(timerelmp, followup) ~ 1, data = df2)
fit.KM2
```


```{r}
ggsurvplot(fit = fit.KM2, surv.median.line = "hv", ylab ="Percent followed")
```


and for survival analyses with two groups:
```{r}
fit.KM3 <- survfit(Surv(timerelmp, Status) ~ groups, data = df2)
fit.KM3
```

```{r}
summary(fit.KM3)
```

```{r}
ggsurvplot(fit = fit.KM3, pval = T, conf.int = T)
```
### Life table method

## Alternative observation schemes

### Death times of elderly residents of a retirement community

1. You need to be at a certain age to be retired, and be alive at the given interval therefore the data is left-truncated. 

2. We don't know how many people joined to the study during the 5 year period, therefore we can't infer the actual death rate.

